A Formal Model of Checked C

### 2023-01-16 sandp23/evaluation.tex
Arun- Implementation Section

\myparagraph{Observations}
Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
\begin{itemize}
  \item Run-time Overhead is proportional to the extent of 
annotated pointers and sandboxed code.
  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
  \item Marshalling data might sometimes be cost-effective instead of propagating the taintedness of a pointer
  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
  \item Conversion style and extent qualitatively dictate the converted program's conversion efforts and is modest at a couple of days. 
\end{itemize}

>>>>

\myparagraph{Observations}
Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
\begin{itemize}
  \item Run-time Overhead is proportional to the extent of 
annotated pointers and sandboxed code.
  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
\end{itemize}

### 2023-01-16 sandp23/implementation.tex
Arun- Implementation Section

\noindent\emph{\_Callback}:
Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\sect{}).

\noindent\emph{\_Mirror}: 
This qualifier copies the corresponding function both into~\cregion and~\ucregion code. Such that callers from each region will invoke the corresponding copy.
This helps handle certain simple utility functions that might be called from both~\cregion and~\ucregion code.
\aravind{Give an example from our evaluation program. (E.g., function foo in program1)}.

@@

Developers can use the~\code{_Mirror} qualifier, which copies the function into both the region so that callers in each region will invoke the corresponding copy avoiding unnecessary tainting.

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions avoiding unnecessary copying of tainted data.
Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.

For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then invoke~\code{strncat(dst, tmp, n);}.
Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
The~\code{_TLIB} qualifier enables this as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// Put your example annotation here.
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the above _TLIB annotation.
strncat(dst, t_str, n);
\end{minted}
\aravind{Arun, Fix the above code. i.e., Add \_TLIB annotation to strncpy }
The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.
Note that developer needs to manually add the check before calling~\code{strcat}.

@@

\iffalse
\noindent\emph{Handling Tainted Buffers in~\cregion}:
All buffers that are marked as tainted in~\cregion should be allocated in the sandbox. For instance,~\code{buff} in function~\code{handle_request} of~\lst{lst:final}.
We perform source rewriting and make the buffer into a pointer variable and allocate the corresponding number of bytes using the sandbox-specific allocator. For instance, we perform the following rewriting for~\code{buff} in~\lst{lst:final} with WASM sandbox.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
- char buff[MAX_MSG_SIZE] __Tainted;
+ char *buff = ....
\end{minted}
\aravind{Arun, please fix the above example on how exactly we replace it with WASM allocator.}
Note that we also add the necessary deallocator calls at function exit points.
\fi

\noindent\emph{Handling Calls to Tainted Functions}:
In~\cregion, we also need to modify calls to tainted functions as they execute inside the sandbox (separate address space) and thus cannot be invoked as regular functions.
However, modifying every call site of tainted functions is tedious and also requires precise pointer analysis~\cite{milanova2002precise} to handle indirect calls through function pointers.

We handle this by~\emph{indirection}: Instead of modifying the call sites, we modify the body of tainted functions to invoke the corresponding function in the sandbox.
For instance, we modify the body of tainted function~\code{process_req1} (\lst{lst:final}) in~\cregion as below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
int process_req1(char *msg, size_t m_l) _Tainted {
- int rc = -1, i;
- if (m_l > MIN_SIZE) {
- ...
+ HOW DO WE HANDLE SANDBOX CALL.
}
\end{minted}
This ensures that all calls (even through function pointers) to the tainted function~\code{process_req1} are redirected to the sandbox.
\aravind{Arun, add the call on how to call sandbox function.}

\noindent\emph{Handling~\texttt{\_Callback} Qualifiers}:
As mentioned in~\sect{subsubsec:addfuncqual}, functions with these qualifiers can be called from~\ucregion.
Consider the following~\code{StringAuth} function that checks whether the provided user is authenticated and needs to be invoked from~\ucregion and hence a callback.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_Callback _TPtr<char> StringAuth(
               _T_Array_Ptr<const char> usertoken : count(len),
               size_t len) {
...
// Checks whether usertoken is authenticated
/*
 These functions will be restricted to only accept
 tainted parameters.
*/
...
}
\end{minted}
These callback functions are only allowed to use tainted parameters as they will be called from a tainted region.

For each such function, we create a corresponding trampoline function that serves as the entry point for the callback function, as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+ unsigned int _T_StringAuth(unsigned sandbox,
+               unsigned int arg_1,
+               unsigned long int arg_2) {
+   // Perform necessary conversion of arguments.
+   // uname <- conver arg_1
+   // len <- arg_2
+   ret = StringAuth(uname, len);
+   // ret_val <- ret
+   return ret_val;
+ }
\end{minted}
The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs the necessary conversion of pointer arguments, and eventually invokes the callback.

@@

Lastly, we change all instances where the callback function pointer is given to tainted functions to use the handle for the corresponding trampoline a shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
// How do we change the call-sites where the function pointer is passed
// as the call-back.
\end{minted}
\aravind{Arun, please check above code example to see if they make sense.}

\subsubsection{Generating~\ucregion Source Partition}
\label{subsubsec:genucregion}
We create~\ucregion partition by first copying all functions marked as~\code{_Tained} and~\code{_Mirror}.
Next, we convert the partition into regular C by replacing all~\systemname{} tainted types with the corresponding C types.
For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
The~\code{read_msg} function signature in~\lst{lst:final} will be modified as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
size_t read_msg(int sock_fd, char *msg,
-                             size_t sz) _Tainted {
+                             size_t sz) {
\end{minted}

Finally, we convert all calls to~\cregion callbacks with sandbox-specific invocations. For instance, for the~\code{StringAuth} call back (\sect{subsub:gencregion}), we modify~\code{...UNCHECKED FUNC..} as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
SAY HOW WE MODIFY UNCHECKED CODE
TO INVOKE CALLBACK FUNCTIONS.
\end{minted}
\aravind{Arun, please fix the above example}

\subsection{Compiler}
The~\systemname compiler includes the type checker (\sect{}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
We implemented the compiler on top of the existing~\checkedc compiler.

Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
- char buff[MAX_MSG_SIZE] _Tainted;
+ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
+    = sbx_allocate(MAX_MSG_SIZE);
...
+ sbx_free(buff)
return rc;
\end{minted}

For all tainted pointer access, our instrumentation adds a check (executed at runtime) to see we check if the address is within the valid range for the sandbox. An example of the check is shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
_TPtr<int> t = ...
+ if (!is_in_range(t, SBX_LOW(), SBX_HIGH()))
+  handle_violation();
*t = 1;
\end{minted}
We show the added check in C for clarify. However, our instrumentation works on LLVM IR.

Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers.

\subsection{Supporting General Partioning Use Case}
\input{figures/newusecase}

Tainting all functions that are not checked may lead to a large number of transitions to the sandbox, which could add unnecessary overhead. In our example, in~\lst{lst:final}, marking~\code{read_msg} function as tainted would cause the system to transition to and from the sandbox every time a message is read, which could be inefficient.

However, as mentioned in~\sect{subsec:checkedc}, checked pointers can be used alongside regular pointers, so the~\cregion{}~\emph{can} contain regular (unchecked) pointers as well.
%
This allows for flexibility in how the program is partitioned, such as only tainting a few high-risk unchecked functions rather than all of them. In this way, the system can be used as a generic program partitioning technique that balances the trade-off between overhead and security assurance.

For our example in~\lst{lst:final}, we can make only the function \code{process_req1} as tainted, thereby transitioning into a sandbox only when the corresponding (high-risk) request is received.

The general guidelines for this alternative use case are shown in~\fig{fig:mixedmodesupport}.
It is best to annotate all functions with checked types for security purposes. However, if it is difficult to convert all functions, it may be more practical to only mark certain high-risk functions with a high probability of vulnerabilities as tainted. This allows for a flexible partitioning system that balances the need for security assurance with the overhead of converting to~\checkedc.
We add the additional constraint of not allowing casting from taint types to generic types to enable this generic use case without violating the safety guarantees of~\systemname.
Specifically, we disallow the following casting in~\cregion:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
//_t_Ptr<int> y; int *z;
|\faRemove| z = (int *)y; // Not allowed.
\end{minted}
\iffalse

@@

\subsubsection{\textbf{Tainted Structures}}
Tainted Pointers to structures crossing into the sandbox are expected to only allow encapsulation of pointer members that are of the tainted type. Consequently, Tstructs have been introduced as a specialized struct type that enforces compile-time semantics to allow prevent its pointer members from being of non-tainted type. \systemname also restricts tainted pointers to only point to structures of Tstruct type. These semantic restrictions of Tstructs are put in place to prevent checked pointer from leaking into the unchecked regions as members of pointer structures. Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout. 
\fi

>>>>

\noindent\emph{\_Callback}:
Developers should use this qualifier to mark callback functions,~\ie functions in~\cregion, that can be called from the tainted region. 
The~\sourcerewriter inserts appropriate sandbox-dependent mechanisms to enable this (\ref{subsub:gencregion}).

\noindent\emph{\_Mirror}: 
This qualifier makes ~\sourcerewriter copy the corresponding function into both ~\cregion and~\ucregion code such that callers from each region can invoke the corresponding copy.
This helps handle certain simple utility functions that might be called from both ~\cregion and~\ucregion code. For example, ~\code{json_value_get_type} in our evaluation of parsons\_wasm has callers from both the ~\cregion and ~\ucregion. 
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
/* JSON Value API */
_Mirror JSON_Value_Type json_value_get_type  
      (_TPtr<const TJSON_Value> value) {
    return value ? value->type : JSONError;
}

_Mirror _TPtr<TJSON_Object> json_value_get_object
(_TPtr<const TJSON_Value> value) {
    return json_value_get_type(value) == JSONObject ?
    value->value.object : NULL;
}

\end{minted}

@@

Developers can use the ~\code{_Mirror} qualifier, which makes ~\sourcerewriter copy the function into both regions so that callers in each region will invoke the corresponding copy, thereby, avoiding unnecessary tainting.

However, ~\code{_Mirror} qualifier semantically asserts the function to only have callees that are either qualified with ~\code{_Tainted} or ~\code {_Mirror}. For example, ~\code{json_value_get_object}'s only callee 
\newline
~\code{json_value_get_type} is ~\code{\_Mirror} qualified.
Similarly, \systemname enforces ~\code{_Mirror} qualified functions to only access global checked data that are marked constant. These control-flow and data-flow compile-time semantics are enforced by \systemname to assert region locality and exclusivity amongst qualified function scopes in both the regions.

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions, thereby, avoiding unnecessary copying of tainted data.
Specifically, this annotation enables developers to pass tainted pointers as arguments to library functions in~\cregion.

For instance, consider that in~\cregion, we want to string concat~\code{n} characters from tainted pointer~\code{t_str} to~\code{dst} (a checked pointer).
We cannot use~\code{strncat} because our type-checker restricts this as it expects both the pointers to be non-tainted as we are in~\cregion.
We need to either implement the logic for string concatenation in~\cregion or first copy the contents of~\code{t_str} into a~\cregion temporary buffer (sat~\code{tmp}) and then call~\code{strncpy} as~\code{strncat(dst, tmp, n);}.
Both of these approaches are inefficient and, if not done correctly, might result in security bugs.
If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use~\code{strncat} with tainted source pointer in~\cregion.
Consequently, ~\code{\_TLIB} qualified functions originally implemented in ~\cregion are expected to be non memory-modifying and thereby, its safe use is left at developer's discretion as \systemname does not enforce any semantics for this.  For example, malloc cannot be qualified with ~\code{_TLIB} and thereby is implemented as an indirect call to sandbox-specific allocator.

The~\code{_TLIB} qualifier enables the qualified function to be used as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
/* Append SRC onto DEST.  */
_TLIB extern char *strcat (char *__restrict __dest,
            const char *__restrict __src)
...        
strncat(dst, t_str, n);
\end{minted}
The above qualifier relaxes type-checking for all the arguments to the calls to~\code{strncat}, enabling the developer to use the function freely in~\cregion.

@@

\noindent\emph{Handling Tainted Buffers in~\cregion}:
All buffers that are marked as tainted in~\cregion should be manually allocated in the sandbox. For instance,~\code{buff} in function~\code{handle_request} of~\lst{lst:final}.
We perform source rewriting and make the buffer into a pointer variable and allocate the corresponding number of bytes using the sandbox-specific allocator. For instance, we perform the following rewriting for~\code{buff} in~\lst{lst:final} with WASM sandbox.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
- char buff[MAX_MSG_SIZE] __Tainted;
+ _TPtr<char> buff = (_TPtr<char>)t_malloc(MAX_MSG_SIZE);  
\end{minted}
Note that we also add the necessary deallocator calls at function exit points.

\noindent\emph{Handling Tainted Functions}:
In~\cregion, we also need to modify tainted functions as they need to execute inside the sandbox (separate address space).
Consequently, We rewrite the tainted function's body with its corresponding sandbox invocation, thereby, routing all ~\cregion calls to tainted functions to their respective sandboxed implementations in ~\ucregion.
For instance, we modify the function body of ~\code{process_req1} in~\lst{lst:final} as below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
int process_req1(_T_Array_ptr<char> msg :
count(m_l), size_t m_l)
__tainted {
- int rc = -1, i;
- // complex code
- return rc;
+ return w2c_process_req1(msg, m_l);
}

\end{minted}

\noindent\emph{Handling~\texttt{\_Callback} Qualifiers}:
As mentioned in~\sect{subsubsec:addfuncqual}, functions with these qualifiers can be called from~\ucregion.
Consider the following~\code{StringAuth} function that checks whether the provided user input ~\code{usertoken} is authenticated by accessing checked data. Since this needs to be invoked from ~\ucregion it is annotated as a ~\code{\_Callback}.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_Callback _TPtr<char> StringAuth(
               _T_Array_Ptr<const char> usertoken : count(len),
               size_t len) {
...
// Checks whether usertoken is authenticated
/*
 These functions will be restricted to only accept
 tainted parameters.
*/
...
}
\end{minted}
These callback functions are only allowed to use tainted parameters as they will be called from a tainted region.

For each such function, we create a corresponding trampoline function that serves as the entry point for the callback function, as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
+ unsigned int _T_StringAuth(void* sandbox,
+               unsigned int arg_1,
+               unsigned long int arg_2) {
+   // Perform necessary Type-conversion of arguments.
+   // uname <- conver arg_1
+   // len <- arg_2
+   ret = StringAuth(uname, len);
+   // ret_val <- ret
+   return ret_val;
+ }
\end{minted}

The trampoline function handles the invocations from sandbox (and hence the extra parameter~\code{sandbox}), performs the necessary conversion of pointer arguments, and eventually invokes the callback.

@@

Lastly, we change the tainted function's body to include an indirect call to the sandbox's implementation of the tainted function. However, instead of passing the callback function pointer directly from the argument list, we pass the generated trampoline handle ~\code{__StringAuth__C_} as shown below:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
_Tainted _TPtr<char> StringProc(_TPtr<_TPtr<const char>> user_input,
_TPtr<_TPtr<char>(_TPtr<const char> input, size_t len)>StringAuth) {
-  ...
-  //complex Function Body       
- return StringAuth(one_past_start, string_len);
+ return w2c_StringProc(_SBX_(), (unsigned int)string, __StringAuth__C_);

}
\end{minted}

\subsubsection{Generating~\ucregion Source Partition}
\label{subsubsec:genucregion}
We create~\ucregion partition by first copying all functions and variables marked as~\code{_Tained} and~\code{_Mirror}.
Next, we convert the partition into regular C by replacing all~\systemname{} tainted types with the corresponding C types.
\newline
For instance:~\code{_T_Array_ptr<char>} will be replaced with~\code{char *}.
The~\code{read_msg} function signature in \lst{lst:final} will be modified as shown below in the ~\ucregion:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
- size_t read_msg(int sock_fd, _T_Array_ptr<char> msg,
-                             size_t sz) __Tainted {
+ size_t read_msg(int sock_fd, char* msg,
+                             size_t sz) {
\end{minted}

% ** We Do NOT modify callback function in checked sources **
% Finally, we convert all calls to~\cregion callbacks with sandbox-specific invocations. For instance, for the~\code{StringAuth} call back (\sect{subsub:gencregion}), we modify~\code{...UNCHECKED FUNC..} as shown below:
% \begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{c}
% SAY HOW WE MODIFY UNCHECKED CODE
% TO INVOKE CALLBACK FUNCTIONS.
% \end{minted}
% \aravind{Arun, please fix the above example}

@@

\subsubsection{\textbf{Tainted Structures}}
Tainted Pointers to structures crossing into the sandbox are expected to only allow encapsulation of pointer members that are of the tainted type. Consequently, Tstructs have been introduced as a specialized struct type that enforces compile-time semantics to allow prevent its pointer members from being of non-tainted type. \systemname also restricts tainted pointers to only point to structures of Tstruct type. These semantic restrictions of Tstructs are put in place to prevent checked pointer from leaking into the unchecked regions as members of pointer structures. Tstruct variables follow the conventional layout unless qualified as a \_Decoy variable, in which case, the Tstruct's pointer member accesses follow a 32-Bit layout.
